### 2016-05-10 paperdraft.tex
revisions

\documentclass[iop,apj]{emulateapj}
\pdfoutput=1 %for arxiv submission to use pdf
\usepackage{apjfonts}
\usepackage{natbib}
\usepackage{epstopdf}
\usepackage{gensymb}
\usepackage{longtable}
\usepackage{amsmath,amstext,textcomp}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=magenta]{hyperref} 
\usepackage[all]{hypcap} %Links go to figures. This breaks deluxetables; use \capstartfalse \capstarttrue around deluxetables to fix it.
\renewcommand{\sectionautorefname}{\S} %for \autoref
\renewcommand{\subsectionautorefname}{\S} %for \autoref
\renewcommand{\subsubsectionautorefname}{\S} %for \autoref

\newcommand{\Reff}{R$_{\rm eff}$}
\newcommand{\Msun}{M$_{\odot}$}

%\renewcommand{\deg}{\ensuremath{^{\circ}}\xspace} %defines a degree symbol

\shorttitle{CCGs in the RESOLVE Survey}
\shortauthors{Snyder et al.}

\begin{document}

\title{Compact Core Galaxies in the RESOLVE Survey}
\author{Elaine M. Snyder\altaffilmark{1}}
\author{Sheila J. Kannappan\altaffilmark{1}}
\author{Dara J. Norman\altaffilmark{2}}
\author{Ashley S. Bittner\altaffilmark{1}}
\author{Mark A. Norris\altaffilmark{3}}
\author{Kathleen D. Eckert\altaffilmark{1}}
\author{Ian Dell'Antonio\altaffilmark{4}}
\author{Callie Hood\altaffilmark{1}}
\author{Samantha Dallas\altaffilmark{4}}
\author{David V. Stark\altaffilmark{1}}
\author{Amanda J. Moffett\altaffilmark{1,5}}
\author{the RESOLVE Team}
\affil{$^1$Department of Physics and Astronomy, University of North Carolina, 141 Chapman Hall CB 3255, Chapel Hill, NC 27599, USA; emsnyder@live.unc.edu}
\affil{$^2$National Optical Astronomical Observatory, 950 N Cherry Ave, Tucson, AZ 85719, USA}
\affil{$^3$Jeremiah Horrocks Institute, University of Central Lancashire, Preston, PR1 2HE, United Kingdom}
\affil{$^4$Department of Physics, Brown University, Box 1843, 182 Hope Street, Barus \& Holley, Providence, RI 02912} 
\affil{$^5$International Centre for Radio Astronomy Research (ICRAR), The University of Western Australia, 35 Stirling High-
way, Crawley, WA 6009, Australia}
\begin{abstract}
We analyze the complete set of 96 ``compact core galaxies'' (CCGs) in the volume-limited RESOLVE survey to investigate their formation and evolution in relation to compact galaxies such as ultra compact dwarfs (UCDs), compact ellipticals (cEs), and dwarf ellipticals (dEs) across a broad statistical distribution of environments. To identify CCGs, we use \textsc{Galfit} to perform one- and two-component S\'ersic fits for all galaxies in RESOLVE with UKIDSS $Y$-band imaging (98\% of the survey), which produces seeing-deconvolved effective radius (\Reff) measurements. The one component fits are used to select candidate CCGs with \Reff\ $< 1000$ pc, and we then perform new two component fits to these galaxies with the S\'ersic n of the second component held fixed at 1 for an exponential disk. We finalize the CCG sample by selecting on core-only (first component) \Reff\ $<800$pc, an upper limit that encompasses all of the similarly compact stellar systems (UCDs, cEs, dEs) included in the Archive of Intermediate Mass Stellar Systems (AIMSS) catalog. We quantify the amount of light in the core and envelope of each CCG by taking the logarithm of the ratio of the light in each component. The result is a smooth continuum of CCGs that range from envelope-dominated to core+envelope to core-dominated. With GALEX $NUV$, SDSS $ugriz$, and 2MASS/UKIDSS $YJHK$ data, we derive colors and star formation histories and find that a significant number of CCGs live on the blue sequence and have recently formed stars. We find that CCGs naturally occur in a range of environments from isolated to cluster. We also derive velocity dispersions ($\sigma$) for 8 CCGs from Gemini IFU data and SOAR spectroscopy. Comparing to other RESOLVE galaxies and AIMSS cEs/dEs, we search for CCGs offset to higher or lower dispersion in the dispersion-stellar mass relation, which may indicate tidal stripping or dissipative formation, respectively. Initial results show 6 CCGs likely following the dissipative formation track and 2 CCGs that are slightly offset to higher $\sigma$.

\end{abstract}

\keywords{galaxies: formation, evolution --- surveys}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{intro}

\noindent The term compact stellar system (CSS) encompasses a class of galaxies that spans the radius range between globular clusters (GCs) and normal elliptical galaxies (Es). Different types of CSSs include compact elliptical galaxies \citep[cEs;][]{Faber1973} and ultra compact dwarf galaxies \citep[UCDs;][]{Phillipps2001}. There are many unanswered questions about how CSSs form and evolve through time. One fact that makes the study of these compact galaxies so intriguing is their apparent scarcity, which raises the question as to whether they are an intermediate and short-lived evolutionary phase in the life of a galaxy, or simply hard to find due to their size. Only in the last 15 years have we discovered enough ($\sim 40$) of these objects to enable studies of their formation scenarios.

CSSs have half-light radii (\Reff) ranging typically from $\sim$10--600 pc \citep[e.g.,][]{Norris2014}, and there are multiple ideas for how they form. The smaller of the CSSs, UCDs may simply be the high mass extension of the GC population \citep{Drinkwater2000, Mieske2002}. The is also evidence that UCDs are created via the tidal stripping of nucleated dwarf galaxies \citep{Bekki2001, Bekki2003, Jennings2015, Zhang2015}. \citet{Seth2014} find a supermassive black hole at the center of another UCD, providing further evidence of tidal stripping being common for UCDs. Similarly, cEs are thought to be the result of tidal stripping events, as argued for the prototypical cE, M32 \citep{Choi2002, Graham2002, Huxor2011}. More evidence is provided by the discovery of tidal streams near cEs \citep{SmithCastelli2008a,Chilingarian2009} and one cE, studied in \citet{Kormendy1997}, that is believed to host a central massive black hole. On the other hand, \citet{Wirth1984}, \citet{Kormendy2009}, and \citet{Kormendy2012a} argue that cEs are the low mass extension to the normal E population created via dissipative formation in gas rich environments, such as mergers or in the two-phase formation process detailed in \citet{Oser2010}. One of the most successful recent searches for CSSs is the AIMSS catalog \citep{Norris2014}: with many newly discovered UCDs and cEs to examine, the authors argue that all of the above scenarios for UCD and cE formation occur.

In contrast to CSSs, dwarf elliptical galaxies (dEs) and their nucleated counterparts (dE,Ns) are quite common, especially in clusters where they were first discovered. In fact, \citet{Binggeli1991} find that of 225 galaxies present in the center of Virgo cluster, 174 are classified as dEs or dE,Ns based on their surface brightness profiles. dEs are typically larger than CSSs with radii ranging up to $\sim800$~pc, whereas dE,Ns may have larger envelopes with a core radius <$800$~pc \citep[e.g.,][]{Norris2014}. Similar to cEs, dEs are often thought to be either tidally stripped remnants \citep{Crnojevi2014} or the low mass extension to elliptical galaxies \citep{Kormendy2012a}. They could also be the remnants of late-type disk galaxies or luminous blue compact dwarfs that have been ram-pressure stripped and quenched as they enter the cluster environment \citep{Lisker2013, Crawford2016a}. There is also some evidence that dE,Ns are progenitors of UCDs as they fall deeper into the cluster core and are further stripped \citep{Pfeffer2013,Zhang2015,Liu2015}.

Given the similarities in formation scenarios between CSSs and dEs, we introduce the term ``compact core galaxy'' or ``CCG'' to encompass CSSs and dEs with core radii up to 800~pc. The goal of this paper is to study these galaxies as part of a continuum of compact galaxies in order to differentiate between formation scenarios. Clues to how CCGs form may be given by their environments,  morphology i.e., one- or two-component structure, star formation histories, and kinematics. We highlight these data below.

%environments
\textit{Environments.} CSSs have been found in a large range of environments thus far. UCDs were first discovered in the Fornax cluster \citep{Hilker1999, Drinkwater2000} and have since been found in the cores of several other clusters \citep{Price2009,Madrid2010,Jones2006,Mieske2009,Misgeld2008}, in groups \citep{Evstigneeva2007}, and near field galaxies \citep{Hau2009,Norris2011}. M32 was first discovered in our Local Group \citep{Faber1973} near M31, and more cEs have later been found in clusters \citep{Chilingarian2007,SmithCastelli2012,Price2009} and other groups \citep{Huxor2011,Chilingarian2010}. As stated previously, dEs and dE,Ns are common in clusters \citep{SmithCastelli2012,Koo1994,Guzman1996,Crawford2016a}, but have also been found near groups \citep{Crnojevi2014,Penny2014}. Only cEs have been found ``free floating'', meaning there is no host galaxy at all: \citet{Huxor2013} and \citet{Paudel2014} find cEs separated from other galaxies by $>1$~Mpc.

Environment bears on formation scenarios in different ways. One possibility is that CSSs found in group or cluster environments may be tidally stripped remnants, whereas free floating CSSs may rather be dissipatively formed like Es. However, \citet{Chilingarian2015} argue that these free-floating cEs could actually be tidally-stripped cores that are ejected from groups and clusters during the tidal stripping process. The recent discovery of a free-floating GC and UCD that have likely been tidally stripped in \citet{Sandoval2015} supports this idea.

%structure
\textit{Morphologies.} CSSs have been found to exhibit either one component (i.e., just a core) or two component (i.e., a core plus a surrounding envelope of gas and stars) morphology. \citet{Graham2002} have shown that M32's light profile is best fit with a core plus a surrounding exponential disk profile, which they argue points toward a tidal-stripping origin. \citet{Pfeffer2013} use N-body simulations to predict that UCDs that are tidally stripped remnants will also display multi-component surface brightness profiles due to lingering outer galaxy structures. Conversely, dissipative processes such as the two-phase formation scenario proposed by \citet{Oser2010} cause cores to form as gas rushes to the center and star formation begins. This core is then ``puffed up'' through time as it interacts with other galaxies. The free-floating cE discovered in \citet{Huxor2013}, shown to have only a one-component light profile, may be an example of the first step in this two-phase formation. It's important to note that these effects may vary based on the degree of stripping and how much time has passed since the stripping or dissipative formation occurred.

Since dEs and dE,Ns are already divided into one and two component categories, it's easier to see the two morphologies are commonly found. \citet{Mastropietro2005a} use N-body simulations to show that as the progenitors of dEs enter clusters, their morphologies are greatly disrupted but disks may not be completely destroyed by ram pressure or tidal stripping. The authors go on to  argue that this excludes dEs as being the progenitors of UCDs, however, \citet{Pfeffer2013} seems to dispel this notion.

%colors
\textit{Star formation histories.} 
Colors and star formation rates are useful since they may point to recent star formation as a part of tidal stripping or dissipative formation. The typical thought is that CSSs are ``red and dead'', and some studies even cut on color when selecting CSS samples \citep{Chilingarian2015}. \citet{Drinkwater2000} have examined the stellar populations of CSSs and find that CSSs are usually best fit as having older stellar populations. \citet{Norris2011} find a similar result with one UCD best fit using older stellar populations. \citet{Drinkwater2000} and \citet{Ferrarese2006} find that the cores of dE,Ns tend to have younger stellar populations compared to their host galaxies.

%kinematics
\textit{Kinematics.}  The last useful tool we have for discerning between formation scenarios is the galaxy's stellar velocity dispersion ($\sigma$). \citet{Bender1992} and \citet{Bekki2003} show that $\sigma$ will remain largely unchanged when tidal stripping occurs, even though the \Reff\ and stellar mass will be greatly reduced. Thus, galaxies that are offset to higher $\sigma$ in a Faber-Jackson (stellar mass -- $\sigma$) relation \citep{faber1976} are most likely tidally stripped, while galaxies falling along the relation are most likely the low-mass extension of normal Es.

In this paper, we present a complete sample of CCGs derived from the volume-limited RESOLVE survey, along with key data quantifying their morphologies, environments, star formation histories, and kinematics. We use this sample to examine the formation scenarios and evolution of CSSs and dEs/dE,Ns simultaneously as a spectrum of compact galaxies that span the gap from GCs to Es. Because we have derived a complete sample from RESOLVE, our sample contains CCGs at every evolutionary stage. This paper is organized as follows. We describe our parent sample, the RESOLVE survey, in \autoref{resolve} and our comparison sample, the AIMSS catalog, in \autoref{aimss}. Our methods for performing seeing-deconvolved one and two component fits for the entire RESOLVE sample using \textsc{Galfit} are presented in \autoref{deconv}. We detail the selection of the CCG sample from the RESOLVE survey and quantify our CCG demographics in \autoref{CCGs}. We describe our photometric, environmental, and spectroscopic data and analysis in \autoref{methods}. We present a statistical distribution of environments, star formation histories, and velocity dispersions for the complete sample in \autoref{results}. In \autoref{discussion}, we present a discussion of our results in the context of CCG formation and evolution. Lastly, the summary and future work are given in \autoref{conclusions}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Samples}
\label{samples}

\subsection{The RESOLVE survey}
\label{resolve}

\noindent We use the RESOLVE survey \citep[][Kannappan et al., in prep]{Kannappan2008} as the parent sample for our census of CCGs. RESOLVE is ideal because it is an unusually complete and volume-limited survey, which allows us to find CCGs at all evolutionary stages. RESOLVE falls within the footprint of the Sloan Digital Sky Survey \citep[SDSS,][]{York2000}, and covers equatorial strips in both semesters: A for the northern spring sky and B for the northern fall sky, which overlaps much of Stripe-82. With $\text{H}_{\circ} = 70~\text{km~s}^{-1}$, RESOLVE-A spans $\sim38,400~\text{Mpc}^3$ and occupies the region $131.25\degree~<~\text{RA}~<~236.25\degree$ and $0\degree~<~\text{Dec}~<~5\degree$, while RESOLVE-B spans $\sim13,700~\text{Mpc}^3$ and occupies the region $330\degree~<~\text{RA}~<~45\degree$ and $-1.25\degree~<~\text{Dec}~<~1.25\degree$. Both semesters cover the redshift range 4500 km s$^{-1} <$ cz $< 7000$ km s$^{-1}$.

RESOLVE's exceptional completeness is obtained by making use of multiple redshift surveys to recover galaxies that were missed in SDSS due to (i) fiber collisions, (ii) the ``shredding'' of galaxies in which the pipeline breaks up a single galaxy into individual pieces so no one piece meets the magnitude cut, and (iii) the intentional exclusion of low surface brightness galaxies ($\mu_{50} < 24.5$ mag arcsec$^{-2}$), even if they meet the magnitude cut. We supplement the SDSS main redshift survey with additional redshifts from archival sources, including the Updated Zwicky Catalog \citep{Falco1999}, HyperLeda \citep{Paturel2003}, 6dF \citep{Jones2009}, 2dF \citep{Colless2001}, GAMA \citep{Driver2011}, and ALFALFA \citep{Haynes2011}, to achieve $\sim12\%$ higher completeness in RESOLVE-A and $\sim25\%$ higher completeness in RESOLVE-B \citep{Eckert2015}.

With its improved redshift recovery, RESOLVE is complete down to absolute $r$-band magnitudes of $-17.33$ and $-17.0$ for the A and B semesters, respectively. To derive the stellar mass completeness limits of RESOLVE, \citet{Eckert2016} use the known stellar masses and magnitudes of all RESOLVE galaxies to determine the stellar mass limit above which no more than 2\% of objects have an $r$-band magnitude fainter than the stated magnitude completeness limit. These values are M$_{\star}~\sim~10^{8.9}$~\Msun~for the A semester and M$_{\star}~\sim~10^{8.7}$~\Msun~in the B semester \citep[Figure 8 in ][]{Eckert2016}. These stellar mass completeness limits allow us to reach down to CSSs with sizes and masses similar to cEs (see \autoref{morph} for more details).

An important caveat to RESOLVE's completeness is that the survey is preferentially incomplete for the very objects this study is focused on, cEs. Such round objects can often be mistaken for stars in redshift surveys using tools such as the \texttt{class\_star} parameter in Source Extractor \citep{Bertin1996}. We are engaged in ongoing efforts to use photo-z estimates to recover some of these objects.

\subsection{The AIMSS catalog}
\label{aimss}

\noindent We use the Archive of Intermediate Mass Stellar Systems (AIMSS) catalog \citep{Norris2014,Forbes2014,Janz2015} as a reference sample for our CCGs.  AIMSS is a catalog of spectroscopically confirmed CSSs found near field galaxies and in groups and clusters, and includes both previously known and newly discovered objects. The AIMSS CSS discovery process was as follows. First, a search was conducted in the HyperLeda redshift catalog \citep{Paturel2003} for all galaxies with M$_{\rm B}~<~-15$ and distances between $\sim 7$ and 200 Mpc (close enough to ensure any CSSs with \Reff\ $> 50$ pc would be resolved in any available HST imaging). The team then searched the Hubble Space Telescope archive for WFCP2, ACS, and WFC3 imaging within 150 kpc in projection of the selected galaxies. Next, Source Extractor was run on the HST images, and candidates were identified using a training set of previously known CSSs from the literature. After vetting by eye and cross-matching to ensure none are previously known, spectra were obtained for redshift confirmation and measurements of velocity dispersions. AIMSS also includes literature data for previously known UCDs, cEs, (nucleated) dEs, dwarf spheroidals, dwarf S0s, young massive clusters, and Es. The catalog provides a compilation of absolute $V$-band magnitude, stellar mass, half-light radius, and velocity dispersion for each of its objects. Thus, though not a statistically-defined sample, AIMSS provides a useful reference catalog to which we can compare our CCGs.

\begin{figure*}[b]
\begin{center}
\includegraphics[scale=0.5]{galfitexample.eps}
\caption{Model inputs and outputs from \textsc{Galfit}. The left column shows the UKIDSS $Y$-band image cutout around the galaxy. The middle column shows the best fit \textsc{Galfit} model, and the right column shows the residuals (data-model). The top row shows the one component fit for one of the RESOLVE galaxies, rf0055. As seen in the residual image, the model is overfitting the inner bulge, and underfitting the outer envelope. The bottom row shows the two component fit for the same galaxy. We are now  able to fit both the inner bulge and the outer envelope. The residual is much cleaner, although there may be some spiral structure remaining.}
\label{fig:galfit}
\end{center}
\end{figure*}

\subsection{CCG sample selection}

\subsubsection{Deconvolution of RESOLVE galaxies}
\label{deconv}

\noindent We begin our analysis by deriving seeing-deconvolved \Reff\ measurements for all galaxies in the RESOLVE sample. For this task we employ \textsc{Galfit} \citep{Peng2002} to deconvolve the seeing point spread function (PSF) and to perform one and two component fits to each galaxy. Deconvolution is especially important since the smallest galaxies in RESOLVE are most affected by seeing, and accurate radius measurements are crucial for the study of these objects.

\textsc{Galfit} needs the following input files to function correctly: an initial input file, an image of the galaxy, a PSF image, an image mask, and a constraint file. We obtain or construct these items as follows:

\begin{description}

\item[The initial input file]{The initial input file gives \textsc{Galfit} the file paths to the image, PSF, mask, and constraint file, and also holds the initial guesses for our fit parameters. For the one component fits, we set up our input file to include a section to measure the level of the sky background and another section to fit a S\'ersic profile to the galaxy. The S\'ersic model will find a best fit for the following parameters: central x \& y position of the model, integrated magnitude, \Reff, S\'ersic n, axial ratio (b/a), and position angle. These initial values for these parameters can just be rough guesses, but for our purposes, we use values for the magnitude, \Reff, b/a, and PA that are derived from RESOLVE's custom processed photometric data (see \autoref{phot} for more details).}

\item[Imaging data]{We use publicly-available high resolution $Y$-band imaging from the UKIDSS Large Area Survey \citep{Lawrence2007}. This data set was chosen because it covers both RESOLVE-A and RESOLVE-B and offers $\sim0.8''$ seeing resolution corresponding to $\sim350$pc at RESOLVE distances in turn letting us span down to the sizes of cE-like galaxies. The $Y$ band is advantageous too, since it probes the underlying stellar population and will not skew our measurements to larger radii if the galaxy is currently experiencing a starburst. We download $\sim10'\times10'$ images of each galaxy from the WFCAM Science Archive\footnote{\url{was.roe.ac.uk//}}}. While \textsc{Galfit} does not require images this large to operate normally, we choose this size so that there will be more stars in the field of view which is useful when constructing the PSF image. There are only 25 galaxies in RESOLVE (2\% of the survey) that do not have UKIDSS imaging and thus are excluded from our analysis.

\item[The PSF image]{\textsc{Galfit} uses the PSF image to deconvolve the galaxy light profile from the seeing, thereby removing the effects of atmospheric scattering of the galaxy light as the data was taken. We build our PSF image via the following steps.
\begin{enumerate}

\item We begin by running Source Extractor on the image to identify every object in the frame, including stars, galaxies, and even sometimes hot pixels. Source Extractor outputs a text file that has estimates of each object's flux, elongation, ellipticity, radius, and \texttt{class\_star}.

\item With this information, we select stars to use for building our PSF. The limiting parameters are (1) that the flux must be between $5,000-100,000$ counts to ensure no extremely faint or saturated stars are included, (2) that the ellipticity be less than 0.2 to ensure we are selecting round objects, (3) that the effective radius of the star be greater than 1 pixel to avoid selecting hot pixels, and (4) that \texttt{class\_star} is greater than 0.6 to ensure the objects we're picking are most likely stars. We also make sure the stars are not near the image edges, since there can sometimes be image artifacts there. Lastly, we make sure to not include the galaxy we are going to fit, since the cE-like galaxies can sometimes appear to be stars.

\item We next make $101\times101$ pixel cutouts of each object detected, with the flux peak at pixel $x=51,y=51$.

\item Background subtraction is then performed by taking the median of the entire cutout and then subtracting each pixel by this median value.

\item We then reject any of the cutouts that have a background median value greater than $2\sigma$ from the mean of the background median values for every cutout. This ensures we are not including any cutouts that have stars with large or bright neighboring objects. We are now left with background-subtracted cutouts of stars with no neighbors that could affect our stacking.

\item Before we can combine all the cutouts, we must normalize the flux of each central star to the same amplitude. To do this, we run each cutout through Source Extractor once more to obtain the fluxes of each central star and use these to weight the cutouts accordingly. We calculate the weight for each cutout as follows: (1) we find the central star with the highest flux of all the cutouts and then (2) divide every flux by this maximum flux. We then divide each cutout by its corresponding weight, which makes each central star have the same amplitude.

\item We then take the median of all the weighted cutouts to create a semi-complete PSF.

\item Lastly, we normalize the semi-complete PSF from Step 7 by dividing each pixel in the frame by the total flux in the image. We find the total image flux by taking the mean of each pixel and multiplying that by the number of pixels in the image (so, $101\times101$). This ensures that the final PSF image is normalized, i.e., the sum of the flux in each pixel of the cutout would equal one.

\end{enumerate}

\noindent Our last step is to fit a 2D Gaussian to the PSF to determine a rough estimate seeing full width half max (FWHM), as a sanity check. We find that the average FWHM for all PSFs is $\sim 0.9''$, only slightly worse than the $\sim 0.8''$ seeing expected for UKIDSS imaging.}

\item[The mask image]{\textsc{Galfit} uses the mask image to ignore any objects neighboring our galaxy while it is completing the fit.The first step in making the mask is to create an image that has the same dimensions as our input image but with each pixel value equaling zero. We again use Source Extractor to compile a list of all the objects in the input image, and  use this list to replace the pixel values at the coordinates of the located objects with one.}

\item[The constant file]{The constraint file hosts the limits we are willing to accept for each of the fit parameters. The constraints for x \& y position in the image, integrated magnitude, \Reff, S\'ersic n, axial ratio (b/a), and position angle for the one component fits are listed in \autoref{constrtable}. We change the constraint file for the two component fits later in the analysis.}

\end{description}

\begin{table}[htdp]
\begin{center}
\begin{tabular}{cc} \hline
parameter & constraint  \\ \hline
x \& y position & $\pm5$ pixels from input value \\
integrated magnitude & 0 -- 40 \\ 
\Reff\ & 0.2 -- 750 pixels \\ 
S\'ersic n & 0.1 -- 8 \\ 
axial ratio (b/a) & 0 -- 1 \\
position angle (PA)) & -360 -- 360 \\ \hline
\end{tabular}
\end{center}
\caption{Summary of the constraints imposed on our \textsc{Galfit} models for our initial one component fits for the entire RESOLVE sample.}
\label{constrtable}
\end{table}%

\noindent We use all of the above to obtain a one component, seeing-deconvolved fit for every galaxy in the RESOLVE survey. We then perform the two component fits as follows. We use the output of the one component fit as the basis of the input file for the two component fit. The first two sections in the two component fit initial input file will be identical to the output of the one component fit (except that the names of the files to be output are changed). We then add a third section to include a second S\'ersic fit, and change the initial guess for \Reff\ to be double that of the one component best fit. This gives \textsc{Galfit} an idea to fit on both a core and envelope. The PSF, image, and mask image will all be the same as before. We add to the constraint file the same constraints as listed in \autoref{constrtable}, but for the third section. In other words, there will be two sets of constraints in the file now, with the first set applying to the first S\'ersic fit and the second set applying the second S\'ersic fit. The constraints for each parameter are the same for both sets. We do not further constrain the two component fits since we are looking for general fits to the wide variety of galaxies in RESOLVE and do not want to over constrain our fits.

For both the one and two component fits, \textsc{Galfit} outputs a text file with the best fits for central x \& y position of the model, integrated magnitude, \Reff, S\'ersic n, axial ratio (b/a), and position angle and a $\chi^2/\nu$ as well as a fits file that contains the data, best fit model, and residual images. See \autoref{fig:galfit} for example one and two component fits for a RESOLVE galaxy.

\subsubsection{Selection of CCGs}
\label{CCGs}

\begin{figure*}[hbpt!]
\begin{center}
\includegraphics{CCGpics.eps}
\caption{Five of the 96 CCGs in the final sample arranged in order from envelope-only to envelope+core to core-only. Images are taken from the SDSS-DR8 image list tool. The central image is rf0492, a CCG with possible ram pressure stripping origins (see \autoref{kin}).}
\label{fig:pics}
\end{center}
\end{figure*}

\noindent Now that we have seeing-deconvolved \Reff\ values for the entire RESOLVE sample, we begin our search for compact core galaxies. We use the AIMSS sample to select a maximum cutoff core-only radius for CCGs at 800 pc. This upper limit ensures that we are selecting the most compact galaxies in RESOLVE, but also lets us select a range of CSSs, including cEs and dEs. RESOLVE's lower stellar mass limit of  restricts us from reaching down to UCD masses.

We start our search for CCGs by using the one component fit from the above analysis and select galaxies with \Reff\ values $<1000$ pc. Since we are looking for galaxies with compact cores, going up to \Reff\ $\sim 1000$ pc allows us to select objects with cores that are dominating the light and therefore may have a smaller core \Reff\ in a two component fit. We note that this limit is arbitrary, and will be relaxed in the future (see \autoref{conclusions}). We also limit the  $\chi^2/\nu$ to be $<1.3$, since values above this represent galaxies with large residuals left over, which are normally spiral or irregular galaxies. After this selection, we are left with 124 galaxies in the CCG candidate sample.

Next, we perform new two component fits using different fitting constraints: (1) the central x and y positions are taken from the one component fit and are held constant for both S\'ersic components, (2) the \Reff\ of the second S\'ersic component fit is constrained to always be larger than that of the first S\'ersic component fit, and (3) the S\'ersic n of the second component is held constant at 1 representing an exponentially decreasing disk profile. We let the values of \Reff, integrated magnitude, b/a ratio, and PA change for both fits. By doing this, we are forcing \textsc{Galfit} to fit on a core (the first S\'ersic component fit) and disky envelope (the second S\'ersic component fit) for each CCG candidate.

We then examine the fits by eye and create our final CCG sample by selecting where the core \Reff\ $< 800$ pc and $\chi^2/\nu < 1.3$ to again exclude poor fits. Our sample is thus comprised of 96 CCGs in total (63 from RESOLVE-A and 33 from RESOLVE-B). This represents $\sim6$\% of the total RESOLVE galaxy sample. \autoref{fig:radius} shows their distribution in stellar mass and decomposed \Reff\ compared to other RESOLVE galaxies and the AIMSS cEs and dEs. For the RESOLVE sample, we are plotting the one component fits, while for the CCGs, we are plotting the core \Reff\ from the fixed two component fits.

We quantify the amount of light in the core vs.~the envelope by taking a ratio of the fluxes of the core and the envelope. The flux of each component is computed with $m = -2.5 \times \log({\rm flux})$, where $m$ is the integrated magnitude output from \textsc{Galfit}. For each plot in this paper, we color code the CCGs by the logarithm of the flux ratio. CCGs with more light in the envelope will have log(flux ratio) $<$ 0, and will hereafter be called ``envelope-dominated' CCGs. CCGs with more light in the core will have log(flux ratio) $>$ 0, and will hereafter be called ``core-dominated'' CCGs. CCGs with log(flux ratio) $\sim 0$ will have both a bright core plus an extended envelope, and will be called ``core+envelope'' CCGs. See \autoref{fig:pics} for examples of CCGs on this continuous scale. There are some cases when \textsc{Galfit} cannot find a core or an envelope, and will instead fit an extremely large and faint component instead. Where this happens (5 out of 96 cases), we classify this fit as either all core/all envelope depending on which component was bad (log(flux ratio) = -1 or 1). We also see in 4 out of the 96 cases that the \Reff\ values for each component will be exactly the same value to 3-4 decimal points, indicating a probable bad fit for one component. When this occurs, we classify the galaxy as being all core (log(flux ratio) = 1).

\begin{figure*}[hbpt!]
\begin{center}
\includegraphics{Reff_Mstars.eps}
\caption{We plot here the deconvolved \Reff\ measurements for the entire RESOLVE sample, along with the CSSs of the AIMSS catalog. In grey are the RESOLVE (non-CCG) galaxies, which are plotted using their one component fit \Reff. The CCG sample is using the core \Reff\ of the two component fits, which is why they may appear separated from the full RESOLVE sample. The colors of the CCGs represent the logarithm of the flux ratio between the core and envelope. Galaxies with log(flux ratio) near 0 will have both a core and envelope, while galaxies with log(flux ratio) $\sim \pm 1$ will have more light in the core (+1) or envelope (-1).}
\label{fig:radius}
\end{center}
\end{figure*}

\section{Methods}
\label{methods}

\subsection{Photometry: galaxy colors, stellar and baryonic masses, and star formation rates}
\label{phot}
\noindent The RESOLVE survey overlaps with a number of photometric surveys, including GALEX $NUV$ \citep{Morrissey2007}, SDSS-DR8 $ugriz$ \citep{Aihara2011}, and UKIDSS $YHK$ and/or 2MASS $JHK$ \citep{Skrutskie2006}. These photometry data has been reprocessed to improve upon the standard SDSS data pipeline, as described in detail in \citet{Eckert2015}. The main improvements in this processing are (1) improved sky subtraction following \citet{Blanton2011} for SDSS photometry plus custom background subtraction for UKIDSS/2MASS, (2) definition of elliptical apertures using the sum of the high S/N $gri$ images to determine the PA and b/a ratio yielding improved sensitivity to the outer disk (if present), and (3) application of the same apertures to all bands, which allows us to measure magnitudes for galaxies that may not have been detected by automated survey pipelines, especially low surface brightness galaxies in 2MASS, UKIDSS, and GALEX. Then, total magnitudes are derived for each band using three methods to enable robust estimation of errors: an exponential outer disk fit, a curve of growth, and a color correction  \citep[Figure 2 in][]{Eckert2015}.

Stellar masses, a long-term fractional stellar mass growth rate (FSMGR), and model colors are then derived from these reprocessed $NUVugrizYJHK$ magnitudes using the spectral energy distribution (SED) modeling code described in  \citet{Kannappan2013}. We adopt the model grid from \citet{Kannappan2013} that combines old simple stellar populations with 2-12 Gyr age ranges and young stellar populations with either continuous star formation from 1015 Myr ago to 0-195 Myr ago, or with a simple stellar population with age either 360, 509, 641, 806, or 1015 Myr. A stellar mass is calculated for each model, along with a likelihood. \citet{Kannappan2013} design their model to also estimate FSMGR, which is the ratio of stellar mass created in the last Gyr to the stellar mass previously existing (i.e., an FSMGR of 1 means that the galaxy has doubled its stellar mass in the last Gyr). Likelihood weighted model colors are also output with k-corrections to $z=0$.

Baryonic masses are estimated

%The HI masses for RESOLVE are from the blind 21cm ALFALFA survey \citep{Haynes2011} and from new pointed observations with the GBT and Arecibo telescopes (Stark et al., submitted). We use these masses along with the stellar masses computed in the SED models to estimate baryonic masses, defined as $\text{M}_{bary} = \text{M}_{\star} + 1.4\text{M}_{HI}$.

\subsection{Environment metrics}
\label{env}
\noindent The volume-limited design of RESOLVE is advantageous to this study because it offers excellent environment information, allowing us to search for CCGs in all environments including groups, clusters, walls, and filaments. We explore the environments of CCGs using two different metrics: the mass of their group halo and their distance to another galaxy. The group halo mass allows us to look for CCG-host pairs that are M32--Andromeda analogues where tidal stripping is likely or conversely to look for CCGs in large groups where either ram-pressure or tidal stripping could be important. The distance to neighboring galaxies lets us examine if interaction may occur or look for analogs of the isolated galaxies found by \citet{Huxor2013} and \citet{Paudel2014}.

Galaxy groups are identified using the friends-of-friends algorithm of \citet{Berlind2006} with linking lengths modified as described in \citet{Eckert2016}, and the masses of the group halos are calculated based on the observed total $r$-band luminosity of galaxies in each group using abundance matching, as described in \citet{Blanton2007}. More information on group identification can be found in \citet{Eckert2016}. The nearest 2-dimensional neighbor distances are computed using the RA and Dec coordinates for each galaxy along with a requirement that the neighbor be within $\pm~500$~km~$^{-1}$ of the primary. Three-dimensional distances are also available based on Hubble's Law distances, however the added errors that come with peculiar velocities make these values less reliable.

\subsection{Kinematics}
\label{kindata}

\subsubsection{SOAR spectroscopy}
\noindent The RESOLVE survey gets the bulk of its spectroscopic data from the SOuthern Astrophysical Research (SOAR) Telescope, located at Cerro Tololo, Chile. We employ the Goodman Spectrograph \citep{Clemens2004} along with RESOLVE's custom-built image slicer with three parallel $1''$ slits to observe galaxies in either an emission- or absorption-line kinematic setup depending on whether the stellar continuum or gas emission lines are more prominent. With the absorption-line setup we are able to derive velocity dispersions, which are typically the best kinematic metric for CCGs. We use the highly-efficient 2100 l/mm VPH grating in order to cover a wavelength range 4858-5503\AA, which includes stellar absorption lines such as H$\beta$, the magnesium triplet, and Fe5270/5335. To achieve S/N $\sim$ 25 per \AA, we bin each central spectrum to \Reff.

The reduction process is as follows: we use standard IRAF tasks to perform bias and overscan subtraction, trimming of the science data to remove unnecessary spatial coverage, and flat fielding. We then use \texttt{LA\_COSMIC} \citep{vanDokkum2001} to clean cosmic rays from each science exposure. We next complete a wavelength calibration and transformation, and then perform object tracing and extraction into a 1D spectrum using the IRAF task apall. We lastly combine the individual exposures using the IRAF task scombine.

\subsubsection{Gemini-South spectroscopy}

\noindent A subset of the faintest or smallest galaxies in RESOLVE are observed with the Gemini-South IFU for their kinematics instead of SOAR. For this absorption-line setup, we use the B1200 grating in 1-slit mode to cover a spatial region of $3''\times5''$ and the spectral range 4200-5600\AA. The B1200 grating offers 0.9\AA\ resolution when binned spectrally by 2. We calculate our exposure times such that we achieve S/N $\sim$ 25 per \AA\, after binning by 2 in the spectral direction and summing the fibers out to \Reff.

The spectral reduction proceeds as follows. Using the standard IRAF Gemini-GMOS data packages, we perform bias and overscan subtraction, spatial trimming, flat fielding and fiber identification. We next use the GMOS package gemCRspec, a wrapper for \texttt{LA\_COSMIC}, to clean cosmic rays from each science exposure. We then correct the science frames for quantum efficiency differences between CCDs. Wavelength transformation, sky subtraction, and flux calibration are subsequently performed, and lastly, data cubes are made and summed. We then stack the individual fiber spectra out to \Reff\ to create a 1D spectrum for each galaxy that has S/N $\sim25$. More information on the data reduction for the Gemini GMOS IFU can be found in the Appendix.

\subsubsection{Velocity Dispersion Extraction}

\noindent We measure $\sigma$ from the both the SOAR and Gemini spectra using {\sc pPXF}, the penalized pixel fitting code from \citet{Cappellari2004}. This code fits a combination of high-resolution (FWHM $= 0.55$ \AA) ELODIE-based model spectra from \citet{Maraston2011} to the observed input spectrum, and determines a best fit solution for $\sigma$. In \autoref{fig:ppxffit}, we show an example of a pPXF fit for one of the CCGs observed with the Gemini-South IFU. In black is the observed spectrum that has been summed out to \Reff\ and in red is the pPXF best fit model spectrum. A caveat to these fits is that they are not corrected for any stellar rotation, which will make absorption lines wider and thus skew the derived $\sigma$ values toward higher values depending on the amount of rotation in each galaxy. Future work will include implementing a stellar rotation subtraction routine for these fits, but preliminary testing by binning the spectra to $1''$ instead of to \Reff\ shows little change between the derived $\sigma$ values.

\begin{figure}[hbpt!]
\begin{center}
\includegraphics[scale=0.4]{rf0078ppxffit.eps}
\caption{Summed Gemini-South IFU spectrum for the CCG rf0078. Overplotted in red is the pPXF best fit model spectrum. The spectrum is zoomed-in to show the MgB triplet fit at $\sim 5275$ \AA.}
\label{fig:ppxffit}
\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{results}

\subsection{Morphologies}
\label{morph}

We plot in \autoref{fig:radius} the deconvolved \Reff\ measurements for the entire RESOLVE sample, along with the CSSs and dEs of the AIMSS catalog. In grey are all RESOLVE galaxies, plotted using their one component fit \Reff. The CCG sample is plotted again with the core \Reff\ of the two component fits, causing them to shift down from the full RESOLVE sample. The symbol color for the CCGs represents the logarithm of the light ratio between the core and envelope, meaning that galaxies with log(flux ratio) near 0 will have both a core and envelope, while galaxies with log(flux ratio) $\sim \pm 1$ will have more light in the core (+1) or envelope (-1). We see in \autoref{fig:radius} that both core- and envelope-dominated CCGs tend to be at the higher end of the \Reff\ values, while core+envelope CCGs tend to be the lower \Reff\ values. There also are more core-dominated CCGs at the higher M$_{\star}$ end while envelope-dominated CCGs tend to reside at low M$_{\star}$.

\subsection{Colors and Star formation rates}

\begin{figure*}[hbpt!]
\begin{center}
\includegraphics[scale=0.65]{sfr_mbary.eps}
\caption{\textbf{(a.)} On the left we plot the $u-r$ colors against stellar mass of CCGs with the entire RESOLVE sample. \textbf{(b.)} On the right we have plotted the FSMGR against baryonic mass for both the CCGs and entire RESOLVE sample. }
\label{fig:fsmgr}
\end{center}
\end{figure*}

\noindent We begin by examining the colors of the CCG sample. The left side of \autoref{fig:fsmgr} shows the k-corrected and extinction-corrected $u-r$ colors (as described in \autoref{phot}) of our CCG sample plotted against M$_{\star}$, compared to the RESOLVE sample as a whole. An unusual finding is that many CCGs live on the blue sequence, since most CSSs are typically thought of as being ``red and dead'', except for dE,Ns with blue cores. The right side of \autoref{fig:fsmgr} shows FSMGR vs. M$_{\rm bary}$ for the same objects. As with the colors, we find that many CCGs are forming stars at a rapid rate (FSMGR $\sim1$ means that the galaxy has doubled its mass in the last Gyr), again an unexpected finding which could indicate that some CCGs are still in the process of forming (if dissipatively formed) or relaxing (if tidally-stripped remnants).

Looking now at the log(flux ratio values), we see that red sequence CCGs tend to be more core-dominated, while blue sequence CCGs tend to me more envelope-dominated. There also tends to be more core-dominated CCGs at lower FSMGR, while envelope-dominated CCGs reside at higher FSMGR. There is considerable scatter though, making these trends harder to qualify.

\subsection{Environments}
\noindent We investigate the distribution of CCGs throughout the cosmic web, and find that CCGs span a range of different environments (see \autoref{fig:envplot}). There are many CCGs residing in low mass groups, similar to the M32--Andromeda relationship, which suggests that tidal stripping could have occurred for those CCGs. On the opposite end, we see several CCGs in massive group halos, and these CCGs tend to be either core- or envelope dominated. This hints towards ram pressure stripping being at play in these environments, especially for those CCGs at larger distances from other galaxies. There are also nine CCGs that are $>1$ Mpc away from another object in RA--Dec--redshift space, and four of those are core dominated, resembling the isolated cE found by \citet{Huxor2013}. These free floating CCGs may be the result of dissipative formation since there is no nearby object to have performed the stripping or may have been ejected from groups or clusters during the stripping process.

\begin{figure*}[hbpt!]
\begin{center}
\includegraphics{nn_groupmass_2d.eps}
\caption{We plot here the group halo mass against the nearest 3D neighbor for the RESOLVE survey and CCG sample. We see that CCGs live in a variety of environments, from extremely isolated (nearest neighbor $>$ 1 Mpc) to massive (M$_{\rm halo} \sim 10^{13}$ M$_{\odot}$) groups. There is no obvious trend into whether cores+envelopes and cores/envelopes preferentially live in different environments.}
\label{fig:envplot}
\end{center}
\end{figure*}

\subsection{Kinematics}
\label{kin}
\noindent We lastly investigate the kinematics of the CCG sample using the $\sigma$s derived from SOAR and Gemini telescope data. \autoref{fig:sigma} plots the Faber-Jackson ($\sigma$-M$_{\star}$) relation for the 8 CCGs with $\sigma$ measurements along with the other 80+ RESOLVE galaxies with $\sigma$ measured. Galaxies with $\sigma$ offset to higher values point to tidal stripping occurring, since they will have less stellar mass than the motions of the stars suggest should be there. Conversely, galaxies with $\sigma$ that lie on the trend may point dissipative formation occurring, since these will have formed similarly to normal elliptical galaxies. Most CCGs seem to be following the dissipative merger track, and in \autoref{discussion}, we examine in depth the formation scenarios for each of these 8 CCGs with velocity dispersion measurements.

\begin{figure*}[b]
\begin{center}
\includegraphics{faber-jackson_resolvesigmas.eps}
\caption{The Faber-Jackson relation (stellar mass vs. velocity dispersion) for the RESOLVE sample, CCGs, and the CSSs of the AIMSS catalog. We look for CCGs offset to higher $\sigma$ for its stellar mass, which may indicate tidal stripping as a possible formation scenario. The green point is M32, the prototypical cE, that is believed to be tidally stripped \citep{Faber1973}. In orange is the free floating cE discovered in \citet{Huxor2013} that is thought to be dissipatively formed.}
\label{fig:sigma}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CCG formation processes} %kind of like a discussion section
\label{discussion}

\noindent We now take an in depth look at the 8 CCGs with kinematics, and assign formation scenarios based on their core \Reff\, morphology, star formation histories, kinematics, and environments. For this discussion, we classify CCGs using the terms ``CSS-like'', ``dE-like'', or ``crossover''.  A CSS-like CCG will have a small core radius and a $\sigma$ value offset higher off the main Faber-Jackson relation, like is expected for UCDs and cEs that have been tidally stripped. A dE-like CCG will have a large core radius and low $\sigma$ like is expected for dEs that are dissipatively formed. Crossover CCGs will have either large core radius and high $\sigma$ or small core radius and low $\sigma$.

We have replotted Figures 3, 5, and 6 in \autoref{fig:miniplots}, this time highlighting the eight CCGs with velocity dispersions only. An important caveat for the CSSs and dEs in the AIMSS catalog is that the radius measurements are not decomposed into one and two components. To correct for this, we plot in the top left panel of \autoref{fig:miniplots} the one component \Reff\ instead of the two component core \Reff\ so that we may accurately compare the radius measurements of CCGs to those of the AIMSS objects.

\subsection{dE-like CCGs}
We classify rs0345, rs1100, rf0003, and rf0328 as dE-like CCGs, based on high one-component effective radius and low $\sigma$. The CCGs rs0345 and rs1100 are both envelope-dominated, live firmly on the blue sequence, and are forming stars quite rapidly with FSMGR~$>1$. Despite having a low stellar mass ($\sim10^{ 8.3}$~\Msun), it's baryonic (gas+stars) mass is higher than many RESOLVE galaxies, suggesting a large gas reservoir fueling star formation. The galaxy is in a moderately-sized group (halo mass $\sim10^{11.7}$~\Msun) and is within $\sim$0.15~Mpc from a companion galaxy. Taken together rs0345 may be an example of tidal stripping or dissipative formation actively occurring, which is fueling the star formation and funneling gas from the disk to build a core.

\subsection{rs1100}
Galaxy rs1100 is envelope-dominated and has one of the highest core \Reff values in the same at $\sim 800$~pc. It is on the blue sequence and has a large FSMGR, suggesting much past star formation. It also lives in a large group with halo mass $\sim10^{12.6}$~\Msun, and has a close neighbor. Its $\sigma$ value places it on the dissipative formation track. We propose that rs1100 may be a galaxy in the process of being ram pressure stripped in the large group environment, or may be dissipatively forming actively.

The CCGs rf0003 and rf0328 are both core-dominated, live intermediate between the blue and red sequences and has a low FSMGR, near $\sim0.10$. It is also in a small group and is $>2$~Mpc from another galaxy, making it perhaps a ``free floating'' CCG, similar to the object found in \citet{Huxor2013}. Its $\sigma$ places it firmly along the main Faber-Jackson relation. All together, these clues point to a dissipatively-formed CCG. However, being near the green valley may indicate that this formation occurred relatively recently.

\ has a higher FSMGR at $\sim0.6$, indicating considerable star formation taking place in the last Gyr. The galaxy resides in a low mass group and is $\sim0.4$~Mpc from a neighbor. Its $\sigma$ places it among the dissipatively formed CCGs, and like rf0003, its color and FSMGR may indicate this formation happened recently.

\subsection{CSS-like CCGs}

\subsection{Crossover CCGs}
We classify rs0567, rs1186, rf0078, rf0492 as crossover CCGs, based on their small radius measurements and low $\sigma$ values. The CCGs rs0567, rs1186, and rf0078 are all three are mildly core-dominated galaxies. They live on the red sequence and display little star formation with FSMGR $> 0.1$ for all. They reside in smaller groups and have moderately close neighbors.

We note that there are CSSs in AIMSS that follow this track with extremely low $\sigma$ (the dark squares at M$_{\star}~\sim~10^{8.5}$~\Msun\ and $\sigma~\sim~20~\text{km~s}^{-2}$). We postulate that these may be low-$z$ analogues to ``red nuggets'' found at high-$z$.

\begin{figure*}[b]
\begin{center}
\includegraphics[scale=0.7]{miniplots.eps}
\caption{We show here \textbf{(a.)} deconvolved \Reff\ vs. stellar mass, \textbf{(b.)} the nearest two-dimensional neighbor vs. group halo mass, \textbf{(c.)} $u-r$ color vs. stellar mass, and \textbf{(d.)} FSMGR vs. baryonic mass for all galaxies in RESOLVE, the cEs and dEs of AIMSS, and the eight CCGs that have velocity dispersion information. This information, combined with \autoref{fig:sigma}, help us distinguish between formation scenarios for the CCG sample.}
\label{fig:miniplots}
\end{center}
\end{figure*}

%\subsection{CCGs without velocity dispersions}

%For the other 88 CCGs without velocity dispersions, we attempt to classify their formation scenarios using the above cases as a primer. We divide the above 8 CCGs into four categories: ``forming with uncertain origins'' (rs0345, rs1100), ``recently formed with dissipative origins likely'' (rf0003, rf0328), ``uncertain stripping and dissipative formation'' (rf0492), and ``likely dissipatively formed'' (rs0567, rs1186, rf0078).

%``Forming with uncertain origins'' CCGs are easily characterized as having envelopes (log(flux ratio) $<$ -0.1), high FSMGR ($>0.5$), and living in high mass groups ($>10^{11.5}$~\Msun). There are 6 CCGs with these parameters ($\sim1\%$ of the sample), including the two already identified. CCGs in the ``likely dissipatively formed'' are also easy characterized by having cores (log(flux ratio) $>$ 0.1), low FSMGR ($>0.5$), and living in low mass groups ($>10^{11.5}$~\Msun). There are 21 CCGs with these parameters ($\sim21\%$ of the sample), again including the three already identified. This leaves 66 CCGs left to classify, but the remaining groups, ``uncertain stripping and dissipative formation'' and ``recently formed'' are more difficult to classify definitely due to their ranges of colors, morphologies, and environments. This leaves $\sim69\%$ of the CCG unclassified until more velocity information is available.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{conclusions}

\noindent In this paper we have examined the colors, FSMGR, environments, and kinematics of the 96 compact core galaxies (CCGs) in the RESOLVE survey. We have selected CCGs to have a seeing-deconvolved, core-only \Reff\ $<$ 800 pc, based on the sizes of the compact and dwarf ellipticals in the AIMSS catalog. We have quantified the amount of light in the cores and envelopes of each CCG by performing two component fits with \textsc{Galfit} and taking the logarithm of the ratio of the fluxes of the two different components. CCGs with more light in the envelope have log(flux ratio) $<$ 0 and are called ``envelope-dominated' CCGs. CCGs with more light in the core have log(flux ratio) $>$ 0 and are called ``core-dominated'' CCGs. CCGs with log(flux ratio) $\sim 0$ have both a bright core plus an extended envelope and are called ``core+envelope'' CCGs. Our main results can be summarized as follows:

\begin{enumerate}
\item \autoref{fig:radius} shows that both core- and envelope-dominated CCGs tend to be at the higher end of the \Reff\ values, while core+envelope CCGs tend to be the lower \Reff\ values. There also are more core-dominated CCGs at the higher M$_{\star}$ end while envelope-dominated CCGs tend to reside at low M$_{\star}$.

\item We find that many CCGs live on the blue sequence and are forming stars at a rapid rate (see \autoref{fig:fsmgr}), which could indicate that some CCGs are still in the process of forming (if dissipatively formed) or relaxing (if a tidally-stripped remnant).

\item We find that CCGs span a range of different environments (see \autoref{fig:envplot}). Many CCGs live in low mass groups, similar to the M32--Andromeda relationship, perhaps hinting at tidal stripping formation scenarios. Others live in massive group halos and tend to be either core- or envelope dominated CCGs, conversely hinting towards ram pressure stripping being at play. There are also many CCGs that are $>1$ Mpc away from another object in RA--Dec--redshift space, resembling the isolated cEs found by \citet{Huxor2013} and \citet{Paudel2014}, which may point to a dissipative formation scenario.

\item Using velocity dispersions ($\sigma$) derived from the SOAR and Gemini telescopes, we classify the eight CCGs with $\sigma$ measurements as ``CCS-like'', ``dE-like'', or ``crossover'', based on their one-component radius measurements (the top left panel of \autoref{fig:miniplots}) and position on the Faber-Jackson ($\sigma$-M$_{\star}$) relation (\autoref{fig:sigma}). We find four CCGs that are dE-like and four that are crossovers.

%In \autoref{fig:sigma}, we find that 6 CCGs tend to follow the main Faber-Jackson trend, which is expected for galaxies formed via dissipative formation, and two CCGs that may be slightly offset to higher $\sigma$.

%\item Using the 8 CCGs with velocity dispersions as a primer, we attempt to quantify the number of CCGs that are tidally stripped or dissipatively formed based on morphologies, colors, FSMGR, and environment. We estimate $\sim$6\% of CCGs are actively forming (though the driver is uncertain) and $\sim$21\% are likely dissipatively formed. The remaining $\sim69$\% need more information to be classified.
\end{enumerate}

\noindent Future work will include the following:

\begin{enumerate}

\item \textit{Redefine CCG sample using $\chi^2$ analysis.}

\item \textit{Investigate gas content.} The amount of neutral Hydrogen gas a CCG has may be able to give us insights into its evolution over time. For example, a CCG in a galaxy cluster with little HI gas may point towards ram pressure stripping being important. Combined with CCG colors and FSMGR, the gas content may be an important tool in determining the evolution of a CCG.

\item \textit{Investigate metallicities.} \citet{Janz2015} have shown that tidally-stripped CSSs are systematically offset to higher metal abundances. With the SOAR data we have in hand, we may be able to test this hypothesis for the CCG sample and use this method as another way to discern formation scenarios.

\end{enumerate}

\noindent \acknowledgments{ \textit{Acknowledgments.}
This work has been supported by funding from NSF grants AST-0955368 and OCI-1156614 and NASA grant HST-AR-12147.01-A. 
Based on observations obtained at the Gemini Observatory under program numbers GS-2013B-Q52, GS-2014B-Q-13, GS-2014B-Q-51, and GS-2015B-C-1, which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina), and Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil).}

\bibliographystyle{apj}
\bibliography{library.bib}

\appendix
\label{app}

\end{document}